---
title: "Dog Adoption Machine Learning"
author: "Michael Martin"
date: "1/12/2020"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(janitor) #For cleaning up names
library(DataExplorer)

```

### Application

###### 
The goal is to predict which independent variables predict a dog being adopted or returned to it's owner.
The dependent variable "placed" is binary indicating whether a dog was either adopted or returned to it's owner or not. This makes it a supervised classification problem, so this is going to be tackled using Random Forest.


```{r include=FALSE}
dogs <- read_csv('G:\\dogs_data_set_final.csv') 
dogs$X1 <- NULL
dogs <- as.tibble(dogs)
```


```{r placed, echo=FALSE}
summary(dogs$placed)
```

This shows that about 76% of dogs in animal shelters are either adopted or returned to their owner. 

```{r include=FALSE}
library(rpart.plot)
library(caTools)
library(rpart)
library(randomForest)

dogs_forest <- dogs %>%
  select(age_years, days_in_shelter, intake_type, placed, intake_condition, out_of_county, fixed.true,
         intimidating_breed_TRUE) %>%
  mutate_if(is.character, as.factor)

#Splitting Data Set into Train and Test data sets 

split <- sample.split(dogs_forest$placed, SplitRatio = .7)
dogs_forest_train = subset(dogs_forest, split == TRUE)

dogs_forest_test = subset(dogs_forest, split == FALSE)

library(randomForest)

set.seed(120)
dog_model_rf <- randomForest(as.factor(placed) ~ .,
                             data = dogs_forest_train,
                             ntree = 1000,
                             mtry = 4,
                             importance = TRUE,
                             replace = FALSE)

dog_model_rf
```

```{r include=FALSE}
library(pROC)
```



```{r}

varImpPlot(dog_model_rf, main = "") %>%
  title(main="Variable Importance Random Forest")
```

#####   Shows the most useful variables for explaining "placed" is probably intake_condition, day_in_shelter, age_years, and intake_type

```{r include=FALSE}
library(verification)

```

```{r}
outcome <- predict(dog_model_rf, dogs_forest_test)
table(dogs_forest_test$placed, outcome, dnn=c("Actual", "Predicted"))

```



```{r}
(489+2242)/(489+263+134+2242)
```

##### This model here states it has about 87% accuracy in the test data set.